Ecosystem carbon balance in the Hawaiian Islands under different scenarios of future climate and land use change

### 2020-07-31 Manuscript/HI_Model_ms.Rmd
updated manuscript files

### *States and transitions*
Transition pathways were modified from @daniel_state-and-transition_2016 and @selmants_baseline_2017. Transitions between state types were pre-defined to represent urbanization, agricultural contraction, agricultural expansion, harvesting of tree plantations, and wildfire. Agriculture, forest, grassland, tree plantation, and shrubland state types each had multiple transition pathways, while the barren state type could only transition to urban. Although several state types could transition to urban, there was no transition pathway from urban to any other state type. Water and wetland state types remained static throughout the simulation period. We developed a high and a low land-use scenario, with transition targets based on historical trends of land use change and population projections [@kim_population_2018].

For the high land-use scenario, annual transition targets for agricultural contraction, agricultural expansion, and urbanization were based on

>>>>

### *States and transitions*
We developed a high and a low land-use scenario, with transition pathways modified from @daniel_state-and-transition_2016. Transitions between state types were pre-defined to represent urbanization, agricultural contraction, agricultural expansion, harvesting of tree plantations, and wildfire. Agriculture, forest, grassland, tree plantation, and shrubland state types each had multiple transition pathways, while the barren state type could only transition to urban. Although several state types had urbanization transition pathways, there was no transition pathway from urban to any other state type. Water and wetland state types remained static throughout the simulation period.

 Transition targets were based on historical trends of land use change and population projections. For the high land-use scenario, rates of agricultural contraction, agricultural expansion, and urbanization were derived from remote sensing-based time series maps of land use change in the Hawaiian Islands from 1992-2011 developed by the National Oceanic and Atmospheric Administration Office of Coastal Planning [@noaa_coastal_2020]. We sampled transition rates for each timestep and Monte Carlo realization from uniform distributions bounded by the median and maximum historical values of agricultural contraction, agricultural expansion, and urbanization for each island. For the low land-use scenario, rates of agricultural contraction and expansion were derived from the same NOAA time series by sampling from uniform distributions bounded by zero and the minimum historical values for each island. Urbanization rates in the low land-use scenario were based on island-level population estimates and projections at five year intervals from 2010-2045 [@kim_population_2018]. We converted population projections into urbanization transition targets following @sleeter_future_2017 by calculating population density for each island and then projecting future developed area based on the five-year incremental change in island population. The spatial extent of agricultural contraction, agricultural expansion, and urbanization was constrained in both land-use scenarios based on existing zoning maps [@daniel_state-and-transition_2016].

The wildfire transition sub-model was modified from @daniel_state-and-transition_2016 using a new wildfire spatial database for the Hawaiian Islands from 1999-2019 (Figure S4). Wildfire size distribution and ignition probabilities for each year were calculated from this 21-year time series for each unique combination of moisture zone (Figure S1), island, and state type (Figure 1). The number and size of fires was randomly drawn from one of these historical year-sets for each timestep and Monte Carlo realization. Wildfire in the low land-use scenario was sampled from the subset of historical fire years at or below the 21-year median area burned statewide, and the high land-use scenario sampled from historical fire years above this 21-year median (Fig. S4a). Burn severity probabilities (high, medium, low) for each combination of moisture zone and state type were taken from @selmants_baseline_2017. Wildfire transitions were restricted to previously burned areas using a map of all wildfires from 1999-2018.
